---
title: "STOR390 Final"
author: "Aarjav Jain"
date: "2024-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**NOTE: The variable name for the SVR method is occasionally misspelled SVVR instead of SSVR. This does not influence the results in any way which still present it as the correct SSVR spelling. It is a harmless typo in parts of the code.**

# Packages

```{r}
install.packages("survivalsvm")
install.packages("survival")
install.packages("survminer")
```

```{r}
library(survivalsvm)
library(survival)
library(survminer)
```

# `veteran` dataset

```{r}
data(veteran, package = "survival")
summary(veteran)
```

```{r}
head(veteran)
```

# SSVR (Support Vector Regression)

```{r}
# Fit the survival SVM model
survsvm_model <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran, 
  type = "regression", 
  gamma.mu = 1, 
  opt.meth = "quadprog", 
  kernel = "lin_kernel"
)

# Print model summary
print(survsvm_model)
```

### Extract predicted risk ranks

```{r}
# Create predictions
predictions <- predict(survsvm_model, newdata = veteran)

# Extract predicted risk ranks
predicted_risks <- as.numeric(predictions$predicted)

length(predicted_risks)
nrow(veteran)
```

### Calculate C Index

```{r}
library(survival)

surv_obj <- Surv(veteran$time, veteran$status)

# Compute C-index
c_index <- concordance(surv_obj ~ predicted_risks)
print(c_index$concordance)
```

### SSVR linear, additive, and radial basis function (RBF) kernels on `veteran`

```{r}
# Linear Kernel
survsvm_linear <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "regression",
  gamma.mu = 1,
  opt.meth = "quadprog",
  kernel = "lin_kernel"
)
predictions_linear <- predict(survsvm_linear, newdata = veteran)
predicted_risks_linear <- as.numeric(predictions_linear$predicted)

# Compute C-index
c_index_linear <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_linear)

# Additive Kernel
survsvm_additive <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "regression",
  gamma.mu = 1,
  opt.meth = "quadprog",
  kernel = "add_kernel"
)
predictions_additive <- predict(survsvm_additive, newdata = veteran)
predicted_risks_additive <- as.numeric(predictions_additive$predicted)

c_index_additive <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_additive)

# RBF Kernel
survsvm_rbf <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "regression",
  gamma.mu = 1,
  opt.meth = "quadprog",
  kernel = "rbf_kernel"
)
predictions_rbf <- predict(survsvm_rbf, newdata = veteran)
predicted_risks_rbf <- as.numeric(predictions_rbf$predicted)

c_index_rbf <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_rbf)

# Print Results
cat("C-index for Linear Kernel:", c_index_linear$concordance, "\n")

cat("C-index for Additive Kernel:", c_index_additive$concordance, "\n")

cat("C-index for RBF Kernel:", c_index_rbf$concordance, "\n")
```

# Ranking (vanBelle1)

### vanbelle1 Linear kernel

```{r}
# vanbelle1 Approach
survsvm_vanbelle1_linear <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "vanbelle1",
  gamma.mu = 1,
  opt.meth = "quadprog",
  kernel = "lin_kernel",
  diff.meth = "makediff1"  # Specify the diff.meth method
)
predictions_vanbelle1_linear_full <- predict(survsvm_vanbelle1_linear, newdata = veteran)
predicted_risks_vanbelle1_linear_full <- as.numeric(predictions_vanbelle1_linear_full$predicted)
c_index_vanbelle1_linear_full <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_vanbelle1_linear_full)

# Print Results
cat("C-index for vanbelle1 Approach:", c_index_vanbelle1_linear$concordance, "\n")
```

### vanbelle1 additive kernel

```{r}
# vanbelle1 Additive Kernel
survsvm_vanbelle1_add <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "vanbelle1",
  gamma.mu = 1,
  opt.meth = "quadprog",
  kernel = "add_kernel",
  diff.meth = "makediff1"  # Specify the diff.meth method
)
predictions_vanbelle1_add_full <- predict(survsvm_vanbelle1_add, newdata = veteran)
predicted_risks_vanbelle1_add_full <- as.numeric(predictions_vanbelle1_add_full$predicted)
c_index_vanbelle1_add_full <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_vanbelle1_add_full)

# Print Results
cat("C-index for vanbelle1 Additive Kernel:", c_index_vanbelle1_add_full$concordance, "\n")
```

### vanbelle1 RBF kernel

```{r}
# vanbelle1 RBF Kernel
survsvm_vanbelle1_rbf <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "vanbelle1",
  gamma.mu = 1,
  opt.meth = "quadprog",
  kernel = "rbf_kernel",
  diff.meth = "makediff1"  # Specify the diff.meth method
)
predictions_vanbelle1_rbf_full <- predict(survsvm_vanbelle1_rbf, newdata = veteran)
predicted_risks_vanbelle1_rbf_full <- as.numeric(predictions_vanbelle1_rbf_full$predicted)
c_index_vanbelle1_rbf_full <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_vanbelle1_rbf_full)

# Print Results
cat("C-index for vanbelle1 RBF Kernel:", c_index_vanbelle1_rbf_full$concordance, "\n")
```

# Hybrid approach

### Hybrid linear kernel

```{r}
# Hybrid Linear Kernel
survsvm_hybrid_linear_full <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "hybrid",
  gamma.mu = c(1, 0.5),  # Two numeric values
  opt.meth = "quadprog",
  kernel = "lin_kernel",
  diff.meth = "makediff1"
)

# Predict on the veteran dataset
predictions_hybrid_linear_full <- predict(survsvm_hybrid_linear_full, newdata = veteran)
predicted_risks_hybrid_linear_full <- as.numeric(predictions_hybrid_linear_full$predicted)

# Compute C-index
c_index_hybrid_linear_full <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_hybrid_linear_full)

# Print Result
cat("C-index for Hybrid Linear Approach:", c_index_hybrid_linear_full$concordance, "\n")
```

### Hybrid additive kernel

```{r}
# Hybrid Additive Kernel
survsvm_hybrid_add_full <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "hybrid",
  gamma.mu = c(1, 0.5),  # Two numeric values
  opt.meth = "quadprog",
  kernel = "add_kernel",
  diff.meth = "makediff1"
)

# Predict on the veteran dataset
predictions_hybrid_add_full <- predict(survsvm_hybrid_add_full, newdata = veteran)
predicted_risks_hybrid_add_full <- as.numeric(predictions_hybrid_add_full$predicted)

# Compute C-index
c_index_hybrid_add_full <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_hybrid_add_full)

# Print Result
cat("C-index for Hybrid Additive Approach:", c_index_hybrid_add_full$concordance, "\n")
```

### Hybrid RBF Kernel

```{r}
# Hybrid RBF Kernel
survsvm_hybrid_rbf_full <- survivalsvm(
  Surv(time, status) ~ .,
  data = veteran,
  type = "hybrid",
  gamma.mu = c(1, 0.5),  # Two numeric values
  opt.meth = "quadprog",
  kernel = "rbf_kernel",
  diff.meth = "makediff1"
)

# Predict on the veteran dataset
predictions_hybrid_rbf_full <- predict(survsvm_hybrid_rbf_full, newdata = veteran)
predicted_risks_hybrid_rbf_full <- as.numeric(predictions_hybrid_rbf_full$predicted)

# Compute C-index
c_index_hybrid_rbf_full <- concordance(Surv(veteran$time, veteran$status) ~ predicted_risks_hybrid_rbf_full)

# Print Result
cat("C-index for Hybrid RBF Approach:", c_index_hybrid_rbf_full$concordance, "\n")
```

# 5-fold cross-validation

### Selecting the same subset for all consistency across all methods and kernels

```{r}
set.seed(123)  # For reproducibility
n_folds <- 5
folds <- sample(rep(1:n_folds, length.out = nrow(veteran)))
```

# SSVR 5-fold

### SSVR Linear 5-fold

```{r}
# Initialize a vector to store C-index values
c_index_values_SVVR_linear <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_SVVR_linear <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "regression",
    gamma.mu = 1,
    opt.meth = "quadprog",
    kernel = "lin_kernel"
  )

  # Predict on the test data
  predictions_SVVR_linear <- predict(model_SVVR_linear, newdata = test_data)
  predicted_risks_SVVR_linear <- as.numeric(predictions_SVVR_linear$predicted)

  # Compute the C-index for the current fold
  c_index_SVVR_linear <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_SVVR_linear)
  c_index_values_SVVR_linear[i] <- c_index_SVVR_linear$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (SVVR Linear):", c_index_SVVR_linear$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_SVVR_linear <- mean(c_index_values_SVVR_linear)
sd_c_index_SVVR_linear <- sd(c_index_values_SVVR_linear)

cat("Mean C-index across 5 folds (SVVR Linear):", mean_c_index_SVVR_linear, "\n")
cat("Standard Deviation of C-index (SVVR Linear):", sd_c_index_SVVR_linear, "\n")
```

### SSVR Additive 5-fold

```{r}
# Initialize a vector to store C-index values for additive kernel
c_index_values_SVVR_add <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_SVVR_add <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "regression",
    gamma.mu = 1,
    opt.meth = "quadprog",
    kernel = "add_kernel"  # Additive kernel
  )

  # Predict on the test data
  predictions_SVVR_add <- predict(model_SVVR_add, newdata = test_data)
  predicted_risks_SVVR_add <- as.numeric(predictions_SVVR_add$predicted)

  # Compute the C-index for the current fold
  c_index_SVVR_add <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_SVVR_add)
  c_index_values_SVVR_add[i] <- c_index_SVVR_add$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (SVVR Additive):", c_index_SVVR_add$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_SVVR_add <- mean(c_index_values_SVVR_add)
sd_c_index_SVVR_add <- sd(c_index_values_SVVR_add)

cat("Mean C-index across 5 folds (SVVR Additive):", mean_c_index_SVVR_add, "\n")
cat("Standard Deviation of C-index (SVVR Additive):", sd_c_index_SVVR_add, "\n")
```

### SSVR RBF 5-fold

```{r}
# Initialize a vector to store C-index values for RBF kernel
c_index_values_SVVR_rbf <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_SVVR_rbf <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "regression",
    gamma.mu = 1,
    opt.meth = "quadprog",
    kernel = "rbf_kernel"  # RBF kernel
  )

  # Predict on the test data
  predictions_SVVR_rbf <- predict(model_SVVR_rbf, newdata = test_data)
  predicted_risks_SVVR_rbf <- as.numeric(predictions_SVVR_rbf$predicted)

  # Compute the C-index for the current fold
  c_index_SVVR_rbf <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_SVVR_rbf)
  c_index_values_SVVR_rbf[i] <- c_index_SVVR_rbf$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (SVVR RBF):", c_index_SVVR_rbf$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_SVVR_rbf <- mean(c_index_values_SVVR_rbf)
sd_c_index_SVVR_rbf <- sd(c_index_values_SVVR_rbf)

cat("Mean C-index across 5 folds (SVVR RBF):", mean_c_index_SVVR_rbf, "\n")
cat("Standard Deviation of C-index (SVVR RBF):", sd_c_index_SVVR_rbf, "\n")
```

# Ranking (vanBelle1) 5-fold

### vanBelle1 Linear 5-fold

```{r}
# Initialize a vector to store C-index values for Vanbelle1 Linear
c_index_values_vanbelle1_linear <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_vanbelle1_linear <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "vanbelle1",  # Vanbelle1 method
    gamma.mu = 1,
    opt.meth = "quadprog",
    kernel = "lin_kernel",  # Linear kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_vanbelle1_linear <- predict(model_vanbelle1_linear, newdata = test_data)
  predicted_risks_vanbelle1_linear <- as.numeric(predictions_vanbelle1_linear$predicted)

  # Compute the C-index for the current fold
c_index_vanbelle1_linear <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_vanbelle1_linear)
  c_index_values_vanbelle1_linear[i] <- c_index_vanbelle1_linear$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Vanbelle1 Linear):", c_index_vanbelle1_linear$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_vanbelle1_linear <- mean(c_index_values_vanbelle1_linear)
sd_c_index_vanbelle1_linear <- sd(c_index_values_vanbelle1_linear)

cat("Mean C-index across 5 folds (Vanbelle1 Linear):", mean_c_index_vanbelle1_linear, "\n")
cat("Standard Deviation of C-index (Vanbelle1 Linear):", sd_c_index_vanbelle1_linear, "\n")
```

### vanBelle1 Additive 5-fold

```{r}
# Initialize a vector to store C-index values for Vanbelle1 Additive
c_index_values_vanbelle1_add <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_vanbelle1_add <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "vanbelle1",  # Vanbelle1 method
    gamma.mu = 1,
    opt.meth = "quadprog",
    kernel = "add_kernel",  # Additive kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_vanbelle1_add <- predict(model_vanbelle1_add, newdata = test_data)
  predicted_risks_vanbelle1_add <- as.numeric(predictions_vanbelle1_add$predicted)

  # Compute the C-index for the current fold
  c_index_vanbelle1_add <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_vanbelle1_add)
  c_index_values_vanbelle1_add[i] <- c_index_vanbelle1_add$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Vanbelle1 Additive):", c_index_vanbelle1_add$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_vanbelle1_add <- mean(c_index_values_vanbelle1_add)
sd_c_index_vanbelle1_add <- sd(c_index_values_vanbelle1_add)

cat("Mean C-index across 5 folds (Vanbelle1 Additive):", mean_c_index_vanbelle1_add, "\n")
cat("Standard Deviation of C-index (Vanbelle1 Additive):", sd_c_index_vanbelle1_add, "\n")
```

### vanBelle1 RBF 5-fold

```{r}
# Initialize a vector to store C-index values for Vanbelle1 RBF
c_index_values_vanbelle1_rbf <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_vanbelle1_rbf <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "vanbelle1",  # Vanbelle1 method
    gamma.mu = 1,
    opt.meth = "quadprog",
    kernel = "rbf_kernel",  # RBF kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_vanbelle1_rbf <- predict(model_vanbelle1_rbf, newdata = test_data)
  predicted_risks_vanbelle1_rbf <- as.numeric(predictions_vanbelle1_rbf$predicted)

  # Compute the C-index for the current fold
  c_index_vanbelle1_rbf <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_vanbelle1_rbf)
  c_index_values_vanbelle1_rbf[i] <- c_index_vanbelle1_rbf$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Vanbelle1 RBF):", c_index_vanbelle1_rbf$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_vanbelle1_rbf <- mean(c_index_values_vanbelle1_rbf)
sd_c_index_vanbelle1_rbf <- sd(c_index_values_vanbelle1_rbf)

cat("Mean C-index across 5 folds (Vanbelle1 RBF):", mean_c_index_vanbelle1_rbf, "\n")
cat("Standard Deviation of C-index (Vanbelle1 RBF):", sd_c_index_vanbelle1_rbf, "\n")
```

# Hybrid 5-fold

### Hybrid Linear 5-fold

```{r}
# Initialize a vector to store C-index values for Hybrid Linear
c_index_values_hybrid_linear <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_hybrid_linear <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "hybrid",  # Hybrid method
    gamma.mu = c(0.5, 0.5),  # Two values for regression and ranking weights
    opt.meth = "quadprog",
    kernel = "lin_kernel",  # Linear kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_hybrid_linear <- predict(model_hybrid_linear, newdata = test_data)
  predicted_risks_hybrid_linear <- as.numeric(predictions_hybrid_linear$predicted)

  # Compute the C-index for the current fold
  c_index_hybrid_linear <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_hybrid_linear)
  c_index_values_hybrid_linear[i] <- c_index_hybrid_linear$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Hybrid Linear):", c_index_hybrid_linear$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_hybrid_linear <- mean(c_index_values_hybrid_linear)
sd_c_index_hybrid_linear <- sd(c_index_values_hybrid_linear)

cat("Mean C-index across 5 folds (Hybrid Linear):", mean_c_index_hybrid_linear, "\n")
cat("Standard Deviation of C-index (Hybrid Linear):", sd_c_index_hybrid_linear, "\n")
```

### Hybrid Additive 5-fold

```{r}
# Initialize a vector to store C-index values for Hybrid Additive
c_index_values_hybrid_add <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_hybrid_add <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "hybrid",  # Hybrid method
    gamma.mu = c(0.5, 0.5),  # Two values for regression and ranking weights
    opt.meth = "quadprog",
    kernel = "add_kernel",  # Additive kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_hybrid_add <- predict(model_hybrid_add, newdata = test_data)
  predicted_risks_hybrid_add <- as.numeric(predictions_hybrid_add$predicted)

  # Compute the C-index for the current fold
  c_index_hybrid_add <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_hybrid_add)
  c_index_values_hybrid_add[i] <- c_index_hybrid_add$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Hybrid Additive):", c_index_hybrid_add$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_hybrid_add <- mean(c_index_values_hybrid_add)
sd_c_index_hybrid_add <- sd(c_index_values_hybrid_add)

cat("Mean C-index across 5 folds (Hybrid Additive):", mean_c_index_hybrid_add, "\n")
cat("Standard Deviation of C-index (Hybrid Additive):", sd_c_index_hybrid_add, "\n")
```

### Hybrid RBF 5-fold

```{r}
# Initialize a vector to store C-index values for Hybrid RBF
c_index_values_hybrid_rbf <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_hybrid_rbf <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "hybrid",  # Hybrid method
    gamma.mu = c(0.5, 0.5),  # Two values for regression and ranking weights
    opt.meth = "quadprog",
    kernel = "rbf_kernel",  # RBF kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_hybrid_rbf <- predict(model_hybrid_rbf, newdata = test_data)
  predicted_risks_hybrid_rbf <- as.numeric(predictions_hybrid_rbf$predicted)

  # Compute the C-index for the current fold
  c_index_hybrid_rbf <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_hybrid_rbf)
  c_index_values_hybrid_rbf[i] <- c_index_hybrid_rbf$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Hybrid RBF):", c_index_hybrid_rbf$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_hybrid_rbf <- mean(c_index_values_hybrid_rbf)
sd_c_index_hybrid_rbf <- sd(c_index_values_hybrid_rbf)

cat("Mean C-index across 5 folds (Hybrid RBF):", mean_c_index_hybrid_rbf, "\n")
cat("Standard Deviation of C-index (Hybrid RBF):", sd_c_index_hybrid_rbf, "\n")
```

# Visualizations

### SSVR Full Dataset

```{r}
# Extract C-index values
c_index_linear_value <- c_index_linear$concordance
c_index_additive_value <- c_index_additive$concordance
c_index_rbf_value <- c_index_rbf$concordanc

# Combine into a dataframe
svvr_results <- data.frame(
  Method = "SSVR",  # Single method
  Kernel = c("linear", "additive", "RBF"),
  `C-index` = c(
    sprintf("%.2f", c_index_linear_value),
    sprintf("%.2f", c_index_additive_value),
    sprintf("%.2f", c_index_rbf_value)
  )
)

# Print the dataframe
print(svvr_results)
```

```{r}
library(kableExtra)

# Generate the table
svvr_results %>%
  kbl(
    caption = "C-index Results for SSVR Method by Kernel",
    col.names = c("Method", "Kernel", "C-index (SD)"),
    align = "lcc",
    booktabs = TRUE
  ) %>%
  kable_styling(full_width = FALSE, position = "center")
```

### vanBelle1 Full Dataset

```{r}
# Extract C-index values for the full dataset
c_index_vanbelle1_linear_full_value <- c_index_vanbelle1_linear_full$concordance
c_index_vanbelle1_add_full_value <- c_index_vanbelle1_add_full$concordance
c_index_vanbelle1_rbf_full_value <- c_index_vanbelle1_rbf_full$concordance

# Combine results into a dataframe
vanbelle1_results_full <- data.frame(
  Method = "vanbelle1",  # Single method
  Kernel = c("linear", "additive", "RBF"),
  `C-index` = c(
    sprintf("%.2f", c_index_vanbelle1_linear_full_value),
    sprintf("%.2f", c_index_vanbelle1_add_full_value),
    sprintf("%.2f", c_index_vanbelle1_rbf_full_value)
  )
)

# Print the dataframe
print(vanbelle1_results_full)

# Generate the table
vanbelle1_results_full %>%
  kbl(
    caption = "C-index Results for Vanbelle1 Method by Kernel (Full Data)",
    col.names = c("Method", "Kernel", "C-index"),
    align = "lcc",
    booktabs = TRUE
  ) %>%
  kable_styling(full_width = FALSE, position = "center")
```

### Hybrid Full Dataset

```{r}
# Extract C-index values for the full dataset
c_index_hybrid_linear_full_value <- c_index_hybrid_linear_full$concordance
c_index_hybrid_add_full_value <- c_index_hybrid_add_full$concordance
c_index_hybrid_rbf_full_value <- c_index_hybrid_rbf_full$concordance

# Combine results into a dataframe
hybrid_results_full <- data.frame(
  Method = "Hybrid",  # Single method
  Kernel = c("linear", "additive", "RBF"),
  `C-index (SD)` = c(
    sprintf("%.2f", c_index_hybrid_linear_full_value),
    sprintf("%.2f", c_index_hybrid_add_full_value),
    sprintf("%.2f", c_index_hybrid_rbf_full_value)
  )
)

# Print the dataframe
print(hybrid_results_full)

# Generate the table
hybrid_results_full %>%
  kbl(
    caption = "C-index Results for Hybrid Method by Kernel (Full Data)",
    col.names = c("Method", "Kernel", "C-index"),
    align = "lcc",
    booktabs = TRUE
  ) %>%
  kable_styling(full_width = FALSE, position = "center")
```

### SSVR 5-fold

```{r}
# Mean and SD for each kernel
mean_c_index_SVVR_linear <- mean(c_index_values_SVVR_linear)
sd_c_index_SVVR_linear <- sd(c_index_values_SVVR_linear)

mean_c_index_SVVR_add <- mean(c_index_values_SVVR_add)
sd_c_index_SVVR_add <- sd(c_index_values_SVVR_add)

mean_c_index_SVVR_rbf <- mean(c_index_values_SVVR_rbf)
sd_c_index_SVVR_rbf <- sd(c_index_values_SVVR_rbf)

# Combine results into a dataframe
svvr_results <- data.frame(
  Method = "SSVR",  # Single method
  Kernel = c("linear", "additive", "RBF"),
  `Mean C-index` = c(mean_c_index_SVVR_linear, mean_c_index_SVVR_add, mean_c_index_SVVR_rbf),
  `SD` = c(sd_c_index_SVVR_linear, sd_c_index_SVVR_add, sd_c_index_SVVR_rbf)
)

# Print the dataframe
print(svvr_results)
```

```{r}
library(kableExtra)

# Generate the table
svvr_results %>%
  kbl(
    caption = "Mean C-index and SD for SSVR Method by Kernel",
    col.names = c("Method", "Kernel", "Mean C-index", "SD"),
    align = "lccc",
    booktabs = TRUE
  ) %>%
  kable_styling(full_width = FALSE, position = "center")
```

### vanBelle 5-fold

```{r}
# Mean and SD for Vanbelle1 Linear Kernel
mean_c_index_vanbelle1_linear <- mean(c_index_values_vanbelle1_linear)
sd_c_index_vanbelle1_linear <- sd(c_index_values_vanbelle1_linear)

# Mean and SD for Vanbelle1 Additive Kernel
mean_c_index_vanbelle1_add <- mean(c_index_values_vanbelle1_add)
sd_c_index_vanbelle1_add <- sd(c_index_values_vanbelle1_add)

# Mean and SD for Vanbelle1 RBF Kernel
mean_c_index_vanbelle1_rbf <- mean(c_index_values_vanbelle1_rbf)
sd_c_index_vanbelle1_rbf <- sd(c_index_values_vanbelle1_rbf)

# Combine results into a dataframe
vanbelle1_results <- data.frame(
  Method = "Vanbelle1",  # Single method
  Kernel = c("linear", "additive", "RBF"),
  `Mean C-index` = c(mean_c_index_vanbelle1_linear, mean_c_index_vanbelle1_add, mean_c_index_vanbelle1_rbf),
  `SD` = c(sd_c_index_vanbelle1_linear, sd_c_index_vanbelle1_add, sd_c_index_vanbelle1_rbf)
)

# Print the dataframe
print(vanbelle1_results)
```

```{r}
# Generate the table
vanbelle1_results %>%
  kbl(
    caption = "Mean C-index and SD for Vanbelle1 Method by Kernel",
    col.names = c("Method", "Kernel", "Mean C-index", "SD"),
    align = "lccc",
    booktabs = TRUE
  ) %>%
  kable_styling(full_width = FALSE, position = "center")
```

### Hybrid 5-fold

```{r}
# Mean and SD for Hybrid Linear Kernel
mean_c_index_hybrid_linear <- mean(c_index_values_hybrid_linear)
sd_c_index_hybrid_linear <- sd(c_index_values_hybrid_linear)

# Mean and SD for Hybrid Additive Kernel
mean_c_index_hybrid_add <- mean(c_index_values_hybrid_add)
sd_c_index_hybrid_add <- sd(c_index_values_hybrid_add)

# Mean and SD for Hybrid RBF Kernel
mean_c_index_hybrid_rbf <- mean(c_index_values_hybrid_rbf)
sd_c_index_hybrid_rbf <- sd(c_index_values_hybrid_rbf)

# Combine results into a dataframe
hybrid_results <- data.frame(
  Method = "Hybrid",  # Single method
  Kernel = c("linear", "additive", "RBF"),
  `Mean C-index` = c(mean_c_index_hybrid_linear, mean_c_index_hybrid_add, mean_c_index_hybrid_rbf),
  `SD` = c(sd_c_index_hybrid_linear, sd_c_index_hybrid_add, sd_c_index_hybrid_rbf)
)

# Print the dataframe
print(hybrid_results)
```

```{r}

# Generate the table
hybrid_results %>%
  kbl(
    caption = "Mean C-index and SD for Hybrid Method by Kernel",
    col.names = c("Method", "Kernel", "Mean C-index", "SD"),
    align = "lccc",
    booktabs = TRUE
  ) %>%
  kable_styling(full_width = FALSE, position = "center")
```

### All methods 5-fold

```{r}
# Combine all methods and their results into a single dataframe
combined_results <- data.frame(
  Method = c(rep("SSVR", 3), rep("Vanbelle1", 3), rep("Hybrid", 3)),
  Kernel = rep(c("linear", "additive", "RBF"), 3),
  `Mean C-index` = c(
    mean_c_index_SVVR_linear, mean_c_index_SVVR_add, mean_c_index_SVVR_rbf,  # SVVR
    mean_c_index_vanbelle1_linear, mean_c_index_vanbelle1_add, mean_c_index_vanbelle1_rbf,  # Vanbelle1
    mean_c_index_hybrid_linear, mean_c_index_hybrid_add, mean_c_index_hybrid_rbf  # Hybrid
  ),
  `SD` = c(
    sd_c_index_SVVR_linear, sd_c_index_SVVR_add, sd_c_index_SVVR_rbf,  # SVVR
    sd_c_index_vanbelle1_linear, sd_c_index_vanbelle1_add, sd_c_index_vanbelle1_rbf,  # Vanbelle1
    sd_c_index_hybrid_linear, sd_c_index_hybrid_add, sd_c_index_hybrid_rbf  # Hybrid
  )
)

# Print the dataframe
print(combined_results)
```

```{r}
library(kableExtra)

# Generate the combined table
combined_results %>%
  kbl(
    caption = "Mean C-index and SD for SVVR, Vanbelle1, and Hybrid Methods by Kernel",
    col.names = c("Method", "Kernel", "Mean C-index", "SD"),
    align = "lccc",
    booktabs = TRUE
  ) %>%
  kable_styling(full_width = FALSE, position = "center")
```

# Novel Analysis

# Gamma hyperparameter tuning for Hybrid

### (0.5, 0.5) - Equal regression and ranking

```{r}
# Initialize a vector to store C-index values for Hybrid RBF
c_index_values_hybrid_rbf_1 <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_hybrid_rbf_1 <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "hybrid",  # Hybrid method
    gamma.mu = c(0.5, 0.5),  # Two values for regression and ranking weights
    opt.meth = "quadprog",
    kernel = "rbf_kernel",  # RBF kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_hybrid_rbf_1 <- predict(model_hybrid_rbf_1, newdata = test_data)
  predicted_risks_hybrid_rbf_1 <- as.numeric(predictions_hybrid_rbf_1$predicted)

  # Compute the C-index for the current fold
  c_index_hybrid_rbf_1 <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_hybrid_rbf_1)
  c_index_values_hybrid_rbf_1[i] <- c_index_hybrid_rbf_1$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Hybrid RBF):", c_index_hybrid_rbf_1$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_hybrid_rbf_1 <- mean(c_index_values_hybrid_rbf_1)
sd_c_index_hybrid_rbf_1 <- sd(c_index_values_hybrid_rbf_1)

cat("Mean C-index across 5 folds (Hybrid RBF):", mean_c_index_hybrid_rbf_1, "\n")
cat("Standard Deviation of C-index (Hybrid RBF):", sd_c_index_hybrid_rbf_1, "\n")
```

### (2, 0.5) - Prioritizes Regression

```{r}
c_index_values_hybrid_rbf_2 <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_hybrid_rbf_2 <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "hybrid",  # Hybrid method
    gamma.mu = c(2, 0.5),  # Two values for regression and ranking weights
    opt.meth = "quadprog",
    kernel = "rbf_kernel",  # RBF kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_hybrid_rbf_2 <- predict(model_hybrid_rbf_2, newdata = test_data)
  predicted_risks_hybrid_rbf_2 <- as.numeric(predictions_hybrid_rbf_2$predicted)

  # Compute the C-index for the current fold
  c_index_hybrid_rbf_2 <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_hybrid_rbf_2)
  c_index_values_hybrid_rbf_2[i] <- c_index_hybrid_rbf_2$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Hybrid RBF):", c_index_hybrid_rbf_2$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_hybrid_rbf_2 <- mean(c_index_values_hybrid_rbf_2)
sd_c_index_hybrid_rbf_2 <- sd(c_index_values_hybrid_rbf_2)

cat("Mean C-index across 5 folds (Hybrid RBF):", mean_c_index_hybrid_rbf_2, "\n")
cat("Standard Deviation of C-index (Hybrid RBF):", sd_c_index_hybrid_rbf_2, "\n")
```

### (0.5, 2) - Prioritizes Ranking

```{r}
c_index_values_hybrid_rbf_3 <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_hybrid_rbf_3 <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "hybrid",  # Hybrid method
    gamma.mu = c(0.5, 2),  # Two values for regression and ranking weights
    opt.meth = "quadprog",
    kernel = "rbf_kernel",  # RBF kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_hybrid_rbf_3 <- predict(model_hybrid_rbf_3, newdata = test_data)
  predicted_risks_hybrid_rbf_3 <- as.numeric(predictions_hybrid_rbf_3$predicted)

  # Compute the C-index for the current fold
  c_index_hybrid_rbf_3 <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_hybrid_rbf_3)
  c_index_values_hybrid_rbf_3[i] <- c_index_hybrid_rbf_3$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Hybrid RBF):", c_index_hybrid_rbf_3$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_hybrid_rbf_3 <- mean(c_index_values_hybrid_rbf_3)
sd_c_index_hybrid_rbf_3 <- sd(c_index_values_hybrid_rbf_3)

cat("Mean C-index across 5 folds (Hybrid RBF):", mean_c_index_hybrid_rbf_3, "\n")
cat("Standard Deviation of C-index (Hybrid RBF):", sd_c_index_hybrid_rbf_3, "\n")
```

### (1, 0.1) - Extreme Regression

```{r}
c_index_values_hybrid_rbf_4 <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_hybrid_rbf_4 <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "hybrid",  # Hybrid method
    gamma.mu = c(1, 0.1),  # Two values for regression and ranking weights
    opt.meth = "quadprog",
    kernel = "rbf_kernel",  # RBF kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_hybrid_rbf_4 <- predict(model_hybrid_rbf_4, newdata = test_data)
  predicted_risks_hybrid_rbf_4 <- as.numeric(predictions_hybrid_rbf_4$predicted)

  # Compute the C-index for the current fold
  c_index_hybrid_rbf_4 <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_hybrid_rbf_4)
  c_index_values_hybrid_rbf_4[i] <- c_index_hybrid_rbf_4$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Hybrid RBF):", c_index_hybrid_rbf_4$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_hybrid_rbf_4 <- mean(c_index_values_hybrid_rbf_4)
sd_c_index_hybrid_rbf_4 <- sd(c_index_values_hybrid_rbf_4)

cat("Mean C-index across 5 folds (Hybrid RBF):", mean_c_index_hybrid_rbf_4, "\n")
cat("Standard Deviation of C-index (Hybrid RBF):", sd_c_index_hybrid_rbf_4, "\n")
```

### (0.1, 1) - Extreme Ranking

```{r}
c_index_values_hybrid_rbf_5 <- numeric(n_folds)

for (i in 1:n_folds) {
  # Split data into training and testing sets using common folds
  train_data <- veteran[folds != i, ]
  test_data <- veteran[folds == i, ]

  # Train the model on the training data
  model_hybrid_rbf_5 <- survivalsvm(
    Surv(time, status) ~ .,
    data = train_data,
    type = "hybrid",  # Hybrid method
    gamma.mu = c(0.1, 1),  # Two values for regression and ranking weights
    opt.meth = "quadprog",
    kernel = "rbf_kernel",  # RBF kernel
    diff.meth = "makediff1"  # Specify difference method
  )

  # Predict on the test data
  predictions_hybrid_rbf_5 <- predict(model_hybrid_rbf_5, newdata = test_data)
  predicted_risks_hybrid_rbf_5 <- as.numeric(predictions_hybrid_rbf_5$predicted)

  # Compute the C-index for the current fold
  c_index_hybrid_rbf_5 <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks_hybrid_rbf_5)
  c_index_values_hybrid_rbf_5[i] <- c_index_hybrid_rbf_5$concordance

  # Print the result for the current fold
  cat("Fold", i, "- C-index (Hybrid RBF):", c_index_hybrid_rbf_5$concordance, "\n")
}

# Evaluate Overall Performance
mean_c_index_hybrid_rbf_5 <- mean(c_index_values_hybrid_rbf_5)
sd_c_index_hybrid_rbf_5 <- sd(c_index_values_hybrid_rbf_5)

cat("Mean C-index across 5 folds (Hybrid RBF):", mean_c_index_hybrid_rbf_5, "\n")
cat("Standard Deviation of C-index (Hybrid RBF):", sd_c_index_hybrid_rbf_5, "\n")
```

# Visualizing Gamma Value Comparisons

```{r}
# Define gamma.mu combinations to test
gamma_values <- list(
  c(0.5, 0.5),
  c(2, 0.5),
  c(0.5, 2),
  c(1, 0.1),
  c(0.1, 1)
)

# Initialize a data frame to store results
results <- data.frame(Gamma1 = numeric(), Gamma2 = numeric(), MeanCIndex = numeric(), StdDevCIndex = numeric())

for (gamma in gamma_values) {
  c_index_values <- numeric(n_folds)
  
  for (i in 1:n_folds) {
    # Split data into training and testing sets
    train_data <- veteran[folds != i, ]
    test_data <- veteran[folds == i, ]
    
    # Train the hybrid model with the current gamma
    model_hybrid <- survivalsvm(
      Surv(time, status) ~ .,
      data = train_data,
      type = "hybrid",
      gamma.mu = gamma,  # Set gamma values
      opt.meth = "quadprog",
      kernel = "rbf_kernel",  # Example with additive kernel
      diff.meth = "makediff1"
    )
    
    # Predict and compute C-index
    predictions <- predict(model_hybrid, newdata = test_data)
    predicted_risks <- as.numeric(predictions$predicted)
    c_index <- concordance(Surv(test_data$time, test_data$status) ~ predicted_risks)
    c_index_values[i] <- c_index$concordance
  }
  
  # Store results for this gamma
  results <- rbind(results, data.frame(
    Gamma1 = gamma[1],
    Gamma2 = gamma[2],
    MeanCIndex = mean(c_index_values),
    StdDevCIndex = sd(c_index_values)
  ))
}

# Print results
print(results)
```

```{r}
ggplot(results, aes(x = as.factor(Gamma1), y = MeanCIndex, color = as.factor(Gamma2))) +
  geom_point(size = 4) +
  scale_color_manual(
    values = c("0.1" = "purple", "0.5" = "blue", "1" = "green", "2" = "red"),
    name = "Ranking Weight (Gamma2)"
  ) +
  labs(
    title = "Gamma Hyperparameter Tuning",
    x = "Regression Weight (Gamma1)",
    y = "Mean C Index Value", 
    color = "Ranking Weight (Gamma2)"
  ) +
  theme_minimal()
```
